A mechanization of semi-simplicial and semi-cubical sets in the calculus of inductive constructions

### 2022-11-25 paper/paper.tex
paper: finish intuitive explanation

% Some abbreviations
\newcommand{\tl}{\ensuremath{\mathsf{tl}}}
\newcommand{\hd}{\ensuremath{\mathsf{hd}}}
\newcommand{\imp}{\rightarrow}
\newcommand{\overright}[1]{\overrightarrow{#1}}
\renewcommand{\D}{D}
\newcommand{\hdD}{D.1}
\newcommand{\tlD}{D.2}
\renewcommand{\d}{d}
\renewcommand{\E}{E}
\newcommand{\ap}{\mathsf{ap}}
\renewcommand{\l}{l}
\renewcommand{\c}{c}
\newcommand{\pair}[2]{#1, #2}
\newcommand{\Type}{\textsf{Type}}
\newcommand{\Dom}{\textsf{Dom}}
\newcommand{\UIP}{\textsf{UIP}}

% These are renewed in tables, but let's define them now for in-document
% typesetting
\newcommandx{\framep}[2][1,2]{\prim{frame}[][#1][#2][][]}
\newcommandx{\layer}[2][1,2]{\prim{layer}[][#1][#2][][]}
\newcommandx{\filler}[2][1,2]{\prim{filler}[][#1][#2][][]}
\newcommandx{\restrf}[3][1,2,3]{\restr{frame}[][#3][][#1][#2][][]}
\newcommandx{\restrl}[3][1,2,3]{\restr{layer}[][#3][][#1][#2][][]}
\newcommandx{\restrc}[3][1,2,3]{\restr{filler}[][#3][][#1][#2][][]}
\newcommandx{\cohf}{\coh{frame}[][][][][][][][]}
\newcommandx{\cohl}{\coh{layer}[][][][][][][][]}
\newcommandx{\cohc}{\coh{filler}[][][][][][][][]}
\newcommandx{\fullframe}[1][1]{\prim{fullframe}[][#1][][][]}

@@

\begin{align*}
  \layer[n][p](D)(d) & \defeq \filler[n-1][p](D.\hd)(\restrf[n][p][L](d)) \times \filler[n-1][p](D.\hd)(\restrf[n][p][R](d))
\end{align*}

We now define $\restrf[n][p][\epsilon](d)$ by recursion on the structure of a frame $d$, which necessitates definitions of $\restrl[n][p][\epsilon](d)(l)$ and $\restrc[n][p][\epsilon](d)(c)$, for $l$ a layer and $c$ a filler. The key case is $\restrc[n][n-1][\epsilon](d)(c)$, where $c$, a $\filler[n][n-1]$, has necessarily the form of $((c_L, c_R), \_)$: $\restrc[n][n-1][L]$ picks out $c_L$, a $\filler[n][n]$, $\restrc[n][n-1][R]$ picks out the $c_R$, also a $\filler[n][n]$, and $\_$ is discarded. There is one more difficulty, which we illustrate by writing down expected and actual types.

@@

\begin{align*}
  c_\omega & : \filler[n-1][p](D.\hd)(\restrf[n-1][p][\omega](d))
\end{align*}

@@

\begin{equation*}
  \filler[n-2][p](D.\hd.\hd)(\restrf[n-1][p][\epsilon](\restrf[n][p][\omega](d)))
\end{equation*}

@@

\begin{equation*}
  \filler[n-2][p](D.\hd.\hd)(\restrf[n-1][p][\omega](\restrf[n][p][\epsilon](d)))
\end{equation*}

Hence, we need a coherence condition to commute the $\restrf[n][p][\epsilon]$ and $\restrf[n][p][\omega]$. Coherence conditions similar to this necessitate, what are shown as, $\cohf$, $\cohl$ and $\cohc$ in tables in the next section.

@@

% Coherence conditions for frame, layer, and filler
\renewcommandx{\cohf}[9][1,2,3,4,5,6,7,8,9]{\coh{frame}[][#2][#3][#4][#5][#6,#7][#8][#9]}
\renewcommandx{\cohl}[9][1,2,3,4,5,6,7,8,9]{\coh{layer}[][#2][#3][#4][#5][#6,#7][#8][#9]}
\renewcommandx{\cohc}[9][1,2,3,4,5,6,7,8,9]{\coh{filler}[][#2][#3][#4][#5][#6,#7][#8][#9]}
\newcommandx{\coht}[9][1,2,3,4,5,6,7,8,9]{\cohtwo{frame}[][#2][#3][#4][#5][#6][#7,#8][#9]}

@@

where $\cohtwo{frame}$ is:

\begin{align*}
  \cohf[m][\omega][\theta][r][p][n][p][][d = {\restrf[m][\epsilon][q+2][n][p][][d = \d, D = \D]}, D = \D]\;\bullet         \\
  \ap (\restrf[m][\omega][r][n][p][][d = {\cohf[m][\epsilon][\theta][q+1][p][n][p][][d = \d, D = \D]}, D = \D]\;\bullet    \\
  \cohf[m][\epsilon][\omega][q][r][n][p][][d = {\restrf[m][\theta][p][n][p][][d = \d, D = \D]}, D = \D] =                  \\
  \ap (\restrf[m][\theta][p][n][p][][d = {\cohf[m][\epsilon][\omega][q+1][r+1][n][p][][d = \d, D = \D]}, D = \D])\;\bullet \\
  \cohf[m][\epsilon][\theta][q][p][n][p][][d = {\restrf[m][\omega][r+1][n][p][][d = \d, D = \D]}, D = \D]\;\bullet         \\
  \ap (\restrf[m][\epsilon][q][n][p][][d = {\cohf[m][\omega][\theta][r][p][n][p][][d = \d, D = \D]}, D = \D])
\end{align*}

\newpage
\section{Future work}

>>>>

% Some abbreviations
\newcommand{\tl}{\ensuremath{\mathsf{tl}}}
\newcommand{\hd}{\ensuremath{\mathsf{hd}}}
\newcommand{\imp}{\rightarrow}
\newcommand{\overright}[1]{\overrightarrow{#1}}
\renewcommand{\D}{D}
\newcommand{\hdD}{D.1}
\newcommand{\tlD}{D.2}
\renewcommand{\d}{d}
\renewcommand{\E}{E}
\newcommand{\ap}{\mathsf{ap}\;}
\renewcommand{\l}{l}
\renewcommand{\c}{c}
\newcommand{\pair}[2]{#1, #2}
\newcommand{\Type}{\textsf{Type}}
\newcommand{\Dom}{\textsf{Dom}}
\newcommand{\UIP}{\textsf{UIP}}

% These are renewed in tables, but let's define them now for in-document
% typesetting
\newcommandx{\framep}[2][1,2]{\prim{frame}[][#1][#2][][]}
\newcommandx{\layer}[2][1,2]{\prim{layer}[][#1][#2][][]}
\newcommandx{\filler}[2][1,2]{\prim{filler}[][#1][#2][][]}
\newcommandx{\restrf}[4][1,2,3,4]{\restr{frame}[][#3][#4][#1][#2][][]}
\newcommandx{\restrl}[4][1,2,3,4]{\restr{layer}[][#3][#4][#1][#2][][]}
\newcommandx{\restrc}[4][1,2,3,4]{\restr{filler}[][#3][#4][#1][#2][][]}
\newcommandx{\cohf}{\coh{frame}[][][][][][][][]}
\newcommandx{\cohl}{\coh{layer}[][][][][][][][]}
\newcommandx{\cohc}{\coh{filler}[][][][][][][][]}
\newcommandx{\coht}{\cohtwo{frame}[][][][][][][][]}
\newcommandx{\fullframe}[1][1]{\prim{fullframe}[][#1][][][]}

@@

\begin{align*}
  \layer[n][p](D)(d) & \defeq \filler[n-1][p](D.\hd)(\restrf[n][p][L][p](d)) \times \filler[n-1][p](D.\hd)(\restrf[n][p][R][p](d))
\end{align*}

$\restrf[n][p][\epsilon][q]$ indicates restriction on $q$ layers of a frame, and the induction is on $p$, going from $0$ to $q$. In particular, $\restrf[n][p][\epsilon][p]$ is a ``full restriction''. We define $\restrf[n][p][\epsilon][p](d)$ by recursion on the structure of a frame $d$, which necessitates definitions of $\restrl[n][p][\epsilon][q](d)(l)$ and $\restrc[n][p][\epsilon][q](d)(c)$, for $l$ a layer and $c$ a filler. The key case is $\restrc[n][p][\epsilon][p](d)(c)$, where $c$, a $\filler[n][p]$, has necessarily the form of $((c_L, c_R), \_)$: $\restrc[n][p][L][p]$ picks out $c_L$, a $\filler[n-1][p]$, $\restrc[n][p][R][p]$ picks out the $c_R$, also a $\filler[n-1][p]$, and $\_$, a $\filler[n][p+1]$, is discarded. There is one more difficulty, which we illustrate by writing down expected and actual types.

@@

\begin{align*}
  c_\omega & : \filler[n-1][p](D.\hd)(\restrf[n-1][p][\omega][q](d))
\end{align*}

@@

\begin{equation*}
  \filler[n-2][p](D.\hd.\hd)(\restrf[n-1][p][\epsilon][q](\restrf[n][p][\omega][p](d)))
\end{equation*}

@@

\begin{equation*}
  \filler[n-2][p](D.\hd.\hd)(\restrf[n-1][p][\omega][p](\restrf[n][p][\epsilon][q+1](d)))
\end{equation*}

Hence, we need a coherence condition to commute the terms. Coherence conditions similar to this necessitate, what are shown as, $\cohf$, $\cohl$ and $\cohc$ in tables in the next section. These are by induction on the structure of $\framep$, $\layer$ and $\filler$. In particular, for $\cohl$, we need a $2$-dimensional coherence condition called $\coht$.

\renewcommandx{\restrf}[7][1,2,3,4,5,6,7]{\restr{frame}[][#2][#3][#4][#5][#6][#7]}
\renewcommandx{\cohf}[9][1,2,3,4,5,6,7,8,9]{\coh{frame}[][#2][#3][#4][#5][#6,#7][#8][#9]}
\renewcommandx{\cohl}[9][1,2,3,4,5,6,7,8,9]{\coh{layer}[][#2][#3][#4][#5][#6,#7][#8][#9]}
\renewcommandx{\cohc}[9][1,2,3,4,5,6,7,8,9]{\coh{filler}[][#2][#3][#4][#5][#6,#7][#8][#9]}
\renewcommandx{\coht}[9][1,2,3,4,5,6,7,8,9]{\cohtwo{frame}[][#2][#3][#4][#5][#6][#7,#8][#9]}

\begin{align*}
  \cohf[m][\omega][\theta][r][p][n][p][][d = {\restrf[m][\epsilon][q+2][n][p][][d = \d]}]\;\bullet     \\
  \ap \restrf[m][\omega][r][n][p][][]\;(\cohf[m][\epsilon][\theta][q+1][p][n][p][][d = \d])\;\bullet   \\
  \cohf[m][\epsilon][\omega][q][r][n][p][][d = {\restrf[m][\theta][p][n][p][][d = \d]}] =              \\
  \ap \restrf[m][\theta][p][n][p][][]\;(\cohf[m][\epsilon][\omega][q+1][r+1][n][p][][d = \d])\;\bullet \\
  \cohf[m][\epsilon][\theta][q][p][n][p][][d = {\restrf[m][\omega][r+1][n][p][][d = \d]}]\;\bullet     \\
  \ap \restrf[m][\epsilon][q][n][p][]\;(\cohf[m][\omega][\theta][r][p][n][p][][d = \d])
\end{align*}

where $\mathsf{ap}$ applies a function on two sides of an equality, and $\bullet$ is transitivity of equality. This property of equality proofs holds in \U.

@@

% Coherence conditions for frame, layer, and filler
\renewcommandx{\cohf}[9][1,2,3,4,5,6,7,8,9]{\coh{frame}[][#2][#3][#4][#5][#6,#7][#8][#9]}
\renewcommandx{\cohl}[9][1,2,3,4,5,6,7,8,9]{\coh{layer}[][#2][#3][#4][#5][#6,#7][#8][#9]}
\renewcommandx{\cohc}[9][1,2,3,4,5,6,7,8,9]{\coh{filler}[][#2][#3][#4][#5][#6,#7][#8][#9]}
\renewcommandx{\coht}[9][1,2,3,4,5,6,7,8,9]{\cohtwo{frame}[][#2][#3][#4][#5][#6][#7,#8][#9]}

@@

\section{Details on the mechanization}
\renewcommandx{\framep}[2][1,2]{\prim{frame}[][#1][#2][][]}
\renewcommandx{\layer}[2][1,2]{\prim{layer}[][#1][#2][][]}
\renewcommandx{\filler}[2][1,2]{\prim{filler}[][#1][#2][][]}
\renewcommandx{\restrf}[4][1,2,3,4]{\restr{frame}[][#3][#4][#1][#2][][]}
\renewcommandx{\restrl}[4][1,2,3,4]{\restr{layer}[][#3][#4][#1][#2][][]}
\renewcommandx{\restrc}[4][1,2,3,4]{\restr{filler}[][#3][#4][#1][#2][][]}
\renewcommandx{\cohf}[1][1]{\coh{frame}[][][][][][][#1][]}
\renewcommandx{\cohl}[1][1]{\coh{layer}[][][][][][][#1][]}
\renewcommandx{\cohc}[1][1]{\coh{filler}[][][][][][][#1][]}

The construction is obviously on induction on $n$, but $\restrf[n]$ depends on $\framep[n]$ and $\framep[n-1]$, while $\cohf[n]$ depends on $\framep[n]$, $\framep[n-1]$, and $\framep[n-2]$. A priori, this necessitates well-founded induction. Due to technical difficulties experienced with Coq, we chose a different route; in practice, since we only need to keep track of three levels, we built separate data structures for the levels, with dependencies.
